---
title: "Exp_analysis"
author: "Caryn Johansen"
date: "April 23, 2015"
output: html_document
---

Get log fold change for gene expression under different nutrient environments
```{r}
library("ggplot2")
library("reshape2")

#shoot_expression <-  read.table("shootnormUQmeangt3.4_range_gt1.5.txt", header=T, sep=" ")
#dim(shoot_expression)
```
#[1] 15528    73

split the data set into four different chunks by P and K nutrient:
```{r}
#shoot_long <- melt(shoot_expression)
#str(shoot_long)
#all the different samples are different factors: 72 factors.

#HPK <- shoot_long[grep("^HPK", shoot_long$variable),]
#LPK <- shoot_long[grep("^LPK", shoot_long$variable),]
#HPLK <- shoot_long[grep("^HPLK", shoot_long$variable),]
#LPHK <- shoot_long[grep("^LPHK", shoot_long$variable),]

#reshape to wide dataframe for analysis
#LPK_wide <- dcast(LPK, Gene ~ variable)
#HPK_wide <- dcast(HPK, Gene ~ variable)
#HPLK_wide <- dcast(HPLK, Gene ~ variable)
#LPHK_wide <- dcast(LPHK, Gene  ~ variable)

#write to table, for future easy of use.
#write.table(LPK_wide, "lpk.txt", sep=",")
#write.table(HPK_wide, "hpk.txt", sep=",")
#write.table(HPLK_wide, "hplk.txt", sep=",")
#write.table(LPHK_wide, "lphk.txt", sep=",")

#now read.table() them in
rm(list=ls())
#LPK <- read.table("lpk.txt", header=T, sep=",")
#HPK <- read.table("hpk.txt", header=T, sep=",")
#LPHK <- read.table("lphk.txt", header=T, sep=",")
#HPLK <- read.table("hplk.txt", header=T, sep=",")
```

Write a function that will get the log fold change between the  
gene expression under the lowest nitrogen condition and all the other  
conditions

```{r}
myShittyFunction <- function( inputDF ) {
  output.df <- data.frame()
  for (i in 1:nrow(inputDF)) {
    #print("row")
    #row <- matrix(nrow=1, ncol=7)
    for (j in 2:length(inputDF)) {
      #print("col")
      datacol <- j-1
      #print(datacol)
      output.df[i,datacol] <- log2(as.numeric(inputDF[i,j])/as.numeric(inputDF[i,2]))
    }
  }
  output.df <- cbind(inputDF[,1], output.df)
  return(output.df)
}

#use said shitty function to get data frames with the LFC
#lfc_LPK <- myShittyFunction(LPK)
#lfc_HPK <- myShittyFunction(HPK)
#lfc_LPHK <- myShittyFunction(LPHK)
#lfc_HPLK <- myShittyFunction(HPLK)

#get the proper column names back
nutrients <- colnames(HPK)
colnames(lfc_HPK) <- nutrients
colnames(lfc_LPK) <- nutrients
colnames(lfc_HPLK) <- nutrients
colnames(lfc_LPHK) <- nutrients
```

If I wanted to plot anything, I'd do this:
```{r}
lfc_HPK_long <- melt(lfc_HPK)
ggplot(lfc_HPK_long, aes(x=variable, y=value, group=Gene)) + geom_point() + geom_line()
```

Histagram plot for the distribution LFC for the genes:
```{r}
hist(lfc_HPK_long$value, breaks=100, freq = F)
```
Nice and normal, as we might expect for a normalized set.

---

I want to know which categorty the maximum and minimum DE is in.
Instead of rank I could just pull out the maximum and the minumum

function to pull out maximum and minimum fold change for each gene
```{r}
anotherShittyFunction <- function( df ) {
  output.m <- matrix(nrow=nrow(df), ncol=4)
  for( i in 1:nrow(df)) {
    
    max <- max(df[i,3:length(df)])
    max.grp <- names(which.max(df[i,3:length(df)]))
    min <- min(df[i,3:length(df)])
    min.grp <- names(which.min(df[i,3:length(df)]))
    
    new.row<-c(max, max.grp, min, min.grp)
    output.m[i,] <- new.row
    
  }
  output.df <- as.data.frame(output.m)
  output.df <- cbind(df[,1], output.df)
  return(output.df)
  
}

hpk_mm <- anotherShittyFunction(lfc_HPK)
```

count, for max and for min, how many of each nitrogen group there are

```{r}
library(plyr)
colnames(hpk_mm) <- c("Gene", "MaxValue", "MaxValueGroup", "MinValue", "MinValueGroup")
hpk_mm$MaxValue <- as.numeric(as.character(hpk_mm$MaxValue))
hpk_mm$MinValue <- as.numeric(as.character(hpk_mm$MinValue))
hpk_mm$MaxValueGroup <- as.character(hpk_mm$MaxValueGroup)

maxGroups <- count(hpk_mm, vars="MaxValueGroup")
minGroups <- count(hpk_mm, vars="MinValueGroup")
ggplot(hpk_mm, aes(x=as.factor(MaxValueGroup))) + geom_bar()
```

Sometimes the maximum value is still a reduction in expression
So I need to split the sets into increase expresion above a threshold
and a decrease expression below a threshold.

For thresholding, let's say 1 and above, for doubling expression
and -1 and below, for halving expression.

```{r}
ExpTrend <- function( df ){
  output.m <- matrix(nrow=nrow(df), ncol=3)
  for (i in 1:nrow(df)) {
    if (max(df[i,3:length(df)]) > 0) {
      de <- max(df[i,3:length(df)])
      trend <- "up"
      de.group <- names(which.max(df[i,3:length(df)]))
    } else {
      de <- min(df[i,3:length(df)])
      trend <- "down"
      de.group <- names(which.min(df[i,3:length(df)]))
    }
    new.row <- c(de, de.group, trend)
    output.m[i,] <- new.row
  }
  output.df <- as.data.frame(output.m)
  output.df <- cbind(df[,1], output.df)
  colnames(output.df) <- c("Gene", "GreatestDE", "DE.Group", "Trend")
  output.df$GreatestDE <- as.numeric(as.character(output.df$GreatestDE))
  return(output.df)
}

count(hpk_trend, var="DE.Group")
count(hpk_trend, var="Trend")

```

Get trends for all nutrient groups:

```{r}
hpk_trend <- ExpTrend(lfc_HPK)
lpk_trend <- ExpTrend(lfc_LPK)
lphk_trend <- ExpTrend(lfc_LPHK)
hplk_trend <- ExpTrend(lfc_HPLK)

count(hpk_trend, var="DE.Group")
count(hpk_trend, var="Trend")
count(lpk_trend, var="DE.Group")
count(lpk_trend, var="Trend")
```

Write all the log fold change and trend tables:
```{r}
#lfc
write.table(lfc_HPK, file="lfc_hpk.txt", sep="\t")
write.table(lfc_LPHK, file="lfc_lphk.txt", sep="\t")
write.table(lfc_LPK, file="lfc_lpk.txt", sep="\t")
write.table(lfc_HPLK, file="lfc_hplk.txt", sep="\t")

#trends
write.table(hpk_trend, file="hpk_trend.txt", sep="\t")
write.table(hplk_trend, file="hplk_trend.txt", sep="\t")
write.table(lpk_trend, file="lpk_trend.txt", sep="\t")
write.table(lphk_trend, file="lphk_trend.txt", sep="\t")
```
